### 2020-03-25 docs/ICTT2019/RMethod_ictt26.tex
minor modifications to ictt text

\begin{abstract}
In this work we apply the Ronen method to resolve the neutron transport in simple homogeneous problems. Slab, cylindrical and spherical geometries are studied. This method demands successive resolutions of the diffusion equation, where the local diffusion constants are modified in order to reproduce new estimates of the currents by a transport operator. The diffusion solver employs here finite differences and the transport-corrected currents are forced in the numerical scheme by means of drift terms, like in the CMFD technique. The solution is compared against reference results provided by a discrete ordinate computer code. Boundary conditions are discussed introducing proper approximations to save the particle balance. Slow convergence is reported on the scalar flux, although the results match the reference solutions in the limit of fine meshes.
%
%In this work we apply the Ronen method in the slab geometry filled with heterogeneous media. Neutron diffusion is solved by a nodal method on a coarse mesh, while reference solutions of the neutron distribution are provided by a discrete ordinate transport solver. The CMFD technique is used for both the implementation of the nodal solver and for the corrections of the diffusion coefficient given by the Ronen method. The presence of discontinuity factors for equivalencing the diffusion solution is investigated jointly with the Ronen method. Results are presented for different levels of material heterogeneity at the slab interfaces.
\end{abstract}
%
% SET RAGGED RIGHT MARGIN
%
\raggedright

@@

This idea was later used by Tomatis and Dall'Osso, who provided a numerical demonstration in a simple slab problem \cite{tomatis2011application}. Instead of updating the diffusion coefficient by the ratio of the current exchange and the flux gradient, as in Fick's law, they recurred to the coarse mesh finite differences method (CMFD) for taking into account the new currents estimated by the integral transport operator in the diffusion solver.
%
%They proposed to update non-linearly the diffusion coefficients at the cell interfaces of the spatial mesh by the coarse mesh finite differences method (CMFD).
%
This technique, largely adopted in the literature of nodal methods \cite{smith1983nodal,lawrence1986progress}, can avoid indeterminate divisions in case of vanishing current and flux gradient. They tested this implementation in a bare slab with two-group cross sections homogenized in a realistic PWR assembly. It was observed that the Ronen method (RM) could drive the flux from diffusion towards the reference of the integral Boltzmann transport equation, regardless of the initial formulation used for the diffusion coefficient. Remarkably, the unphysical zero-flux boundary condition was not preventing to correct the flux near the boundary, that is within a few mean free paths. However, higher errors were noticed at the boundary with vacuum, slowly decreasing even after many iterations.% Their work was limited to the homogeneous slab, and resolved neutron diffusion on the same discretized spatial mesh used for the reference from neutron transport.
%
%In this work, we study the Ronen method with coarse meshes in order to approach the typical situation of core calculations, where homogenized cross sections are prepared in single fuel assemblies by lattice transport calculations. Nodal expansion methods are generally used with coarse meshes relying on polynomial-like or semi-analytical solutions \cite{lawrence1986progress}. They can also employ the discontinuity factors in their solving equations, thus yielding an equivalent form to the reference transport solution \cite{smith1986assembly}. Its implementation through the CMFD compensates the truncation errors caused by the approximate finite differences, and eases the global resolution to many local two-nodes problems. Of course, the simplified 1-D geometry let us disregard here all issues related to the transverse leakage approximation.
%
%Specifically, we compare hereafter the solutions obtained by the Ronen method and by the analytical nodal method, with the aim to present the first method as an on-line equivalence technique, farther capable to fulfill the neutron transport in its integral form on the coarse mesh. %We also use the discontinuity factors to enforce the physical equivalence between diffusion and transport, which is not preserved after the common flux-volume weighting homogenization of the cross sections. The implementation of both methods uses the CMFD, as shown in section \ref{sec:impl}. The basics of the analytical nodal method are reported in section \ref{sec:nd}, while the procedure to estimate the currents by the integral transport operator follows in section \ref{sec:ce}. New transport-corrected boundary conditions are derived in section \ref{sec:bc} to fix the bad performances observed in \cite{tomatis2011application}. The results on a few characteristic configurations are discussed in section \ref{sec:res}. Finally, the article ends with the conclusion in section \ref{sec:cncls}.

@@

The cross sections, as well as the diffusion coefficient, are usually available as volume-averaged data per cell in the mesh. Once the scalar flux is known from the finite differences solver using the original diffusion coefficients, the integral expressions derived in section \ref{sec:ce} can be used to get new estimates of the current $J$ at the cell interfaces. %Alternatively, new current estimates can come from the two-nodes problem resolved by the nodal method in section \ref{sec:nd}.

@@

\subsection{Slab geometry}
\label{sec:slab}
The angular flux from the integral transport equation (with standard notation) is:
\begin{equation}
\varphi(x, \mu) = \varphi(x_b, \mu) e^{-\tau(x_b, x)/\mu} + \int_{x_0}^x {dx' \frac{q(x', \mu)}{\mu} e^{-\tau(x', x)/\mu}},
\label{eq:NTE}
\end{equation}
with the optical length $\tau(x_1, x_2) = \int_{x_1}^{x_2}{dx' \sigma_t(x')}$. Contrary to other formulations, we do not include the direction cosine of the neutron velocity $\mu$ in the definition of the optical length to ease the numerical integrations in the following. Eq. \ref{eq:NTE} holds for both positive directions $(\mu > 0)$ where $x_0 = x_{-1/2} = a$, and negative ones $(\mu > 0)$ where $x_0 = x_{I-1/2} = b$. The inclusion of the energy variable in Eq. \ref{eq:NTE} is straightforward; we use hereafter the multi-group theory, writing the source $q_g$ as\footnote{Curly brackets remind the possible use of the $k$-eigenvalue in case of vanishing external sources.}:
\begin{equation}
\label{eq:srcg}
q_g = \sum_{g'=1}^G{\left[
  \sum_{l=0}^\infty{ \frac{2l+1}{2}\sigma_{s,l, g' \rightarrow g}(x)P_l(\mu)\varphi_{l,g'}(x)
  + \frac{\chi_g}{2\{k\}} \nu\sigma_{f,g'}(x)\phi_{g'}(x)}\right] + S_g
                 },
\end{equation}
where the subscript $l$ refers to the moments of the expansions of the flux and of the scattering cross section on the Legendre polynomials $P_l$. Fission emission is isotropic, so that only the scalar flux $\phi$ being equal to the first moment $\varphi_0$ is retained for the fission term. Without the external source $S$, the multiplication factor $k$ is used as eigenvalue to avoid the trivial vanishing solution. Since $q$ is computed by the results of the diffusion equation in the Ronen method, $l$ can only assume the values 0 and 1.% Hereafter we consider only isotropic sources.

Multiplication of Eq. \ref{eq:NTE} by $\mu$ and the integration over $[-1, 1]$ provides the expression for the current, see \cite{tomatis2011application}. The integration on $\mu$ yields integral exponential functions by a simple change of variable ($\mu = \pm 1/u$, according to the sign of $\mu$) \cite{AS1964handbook}:
\begin{equation*}
E_n(\tau) = \int_0^1{d\mu\, e^{-\tau / \mu} \mu^{n-2}} = \int_1^\infty{du\, e^{-\tau u}u^{-n}}.
\end{equation*}
These special functions are then used to compute the current at the cell interfaces for all $i$:
\begin{equation}
\begin{split}
% J_g(x_{i+1/2}) &=
J_{g,i+1/2} &= \int_0^{ 1}{d\mu\, \mu \varphi_g(a, \mu) e^{-\tau_g(a, x_{i+1/2})/\mu}}
             - \int_0^{-1}{d\mu\, \mu \varphi_g(b, \mu) e^{-\tau_g(b, x_{i+1/2})/\mu}}\\
  & + \sum_{j=0}^{I-1} \left[  \sign(x_{i+1/2} - x_{j+1/2}) \frac{q_{0,g,j}}{2} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{2}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } \right.\\
  & \left. \qquad+ \frac{3}{2}q_{1,g,j} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{3}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) }\right],
\end{split}
\label{eq:J_tr}
\end{equation}
where $q_{(.),g,j}$ are the moments of the souce $q_g$ from Eq. \ref{eq:srcg} averaged in the volume of the $j$-th cell:
\begin{equation*}
\label{eq:srclg}
\begin{split}
   q_{0,g,j} &= \sum_{g'}\left(\sigma_{s,0,g' \rightarrow g,j} +
                              \chi_g \nu\sigma_{f,g',j}
                        \right) \phi_{g',j},\\
   q_{1,g,j} &= \sum_{g'}\sigma_{s,1,g' \rightarrow g,j}J_{g',j}.
\end{split}
\end{equation*}
The current appearing in $q_{1,g,j}$, that is in case of linearly anisotropic scattering, is also to be considered as volume-averaged in the cell $j$. The optical lengths show the subscript $g$ because they are evaluated with the corresponding total cross section $\sigma_{t,g}$. For the property $E'_{n+1}(\ell) = -E_n(\ell)$, the spatial integrals of the integral exponential functions can be solved analytically by integrating on $\tau$:
\begin{equation*}
\begin{aligned}
\int_{x_{j-1/2}}^{x_{j+1/2}}{ dx' E_{n}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } =
\frac{\sign(x_{i+1/2} - x_{j+1/2})}{\sigma_{t,g,j}} \cdot \\
\left[
  E_{n+1} \left(\lvert \tau_g(x_{j+1/2}, x_{i+1/2}) \rvert \right)
 -E_{n+1} \left(\lvert \tau_g(x_{j-1/2}, x_{i+1/2}) \rvert \right)
\right].
\end{aligned}
\end{equation*}
These quantities can be computed for $i > j$ only, because the others are simply opposite in sign.

@@

The projections of partial currents on the outward normal $\hat{n}$ to a given surface are obtained from integration of Eq. \ref{eq:ITE_xy} with the weight $| \hat{n} \cdot \mathbf{\Omega} | = \sin \theta | \hat{n} \cdot \mathbf{\Omega}_p |$ on $\hat{n} \cdot \mathbf{\Omega} \lessgtr 0$. Higher moments are obtained by the weights associated to the corresponding spherical harmonics. Integrations along the polar angle are resolved with the introduction of the Bickley-Naylor functions \cite{amos1983uniform}:
\begin{equation}
\text{Ki}_n(\tau) = \int_0^{\pi/2}{ d\theta \sin^{n-1} \theta \exp\left(-\frac{\tau}{\sin \theta}\right) }.
\label{eq:Kin}
\end{equation}
The properties used for the computation of these functions are available elsewhere \cite{lewis1984computational,hebert2009applied,amos1983algorithm}.

\begin{subequations}
After introducing these arguments for a bare cylinder, the current leaving the cylindrical surface at $\mathbf{r}$ (with direction $\hat{n}$) given by neutrons flying along $\mathbf{\Omega}_p$ is:
\begin{equation}
J^+(\mathbf{r}) = \frac{1}{2\pi}\int_{W^+}{
  d\omega\, | \hat{n} \cdot \mathbf{\Omega}_p | \,
  \int_0^S ds\, \text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r} - s \mathbf{\Omega}_p \right)\right]
  q(\mathbf{r} - s \mathbf{\Omega}_p)},
\label{eq:currp_cylinder1}
\end{equation}
with $W^+ = \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p > 0\right\}$. Using the surface element $dA' = s d\omega ds$ and $\mathbf{r}'_p = \mathbf{r} - s \mathbf{\Omega}_p$, it is also:
\begin{equation}
J^+(\mathbf{r}) = \int_{W^+ \times [0, S]}{
  dA'\, \hat{n} \cdot (\mathbf{r} - \mathbf{r}'_p) \,
  \frac{\text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r}'_p \right)\right]}{2 \pi \lvert \mathbf{r} - \mathbf{r}'_p \rvert^2}
  q(\mathbf{r}'_p)}.
\label{eq:currp_cylinder2}
\end{equation}
\label{eq:currp_cylinder}
\end{subequations}
$J^-$ is obtained with $w \in W^- =  \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p < 0\right\}$ instead.

@@

The corresponding form for the reduced escape probability for the incoming current is $\varepsilon^-_{i+1/2,j}$:
\begin{equation}
\label{eq:esc_prob_cyij-}
  \varepsilon^-_{i+1/2,j} = \frac{2}{r_{i+1/2}}
    \displaystyle ~\int_0^{r_{i+1/2}} { dh\, y
        \int_0^{\ell_j}{
            d\ell \text{Ki}_2 \left(\tau_{ji} + \sigma_j \ell \right)
        }
	},
\end{equation}
for $i<j$, where $\varepsilon^-_{i+1/2,j}\equiv 0$ for $i\geq j$ since in this case a neutron born in volume $j$ (or in volume $j=i$) cannot enter volume $i$ through its outer surface at radius $r_{i+1/2}$ without undergoing a collision.

Singularity of the integrand for the $h$-integration is noticed in the convex part while computing the collision probabilities \cite{hebert2009applied}; this notably happens at the outer radius of $j$-th ring where $\hat{n}$ is perpendicular to $\mathbf{\Omega}_p$. However, that integral can be regularized by a suitable change of variable. % $\varrho = \sqrt{r - h}$.
The calculation of the escape probabilities is not affected by this problem because of the factor $y$ that comes from the cosine of the angle described by the normal $\hat{n}$ and $\mathbf{\Omega}_p$, which cancels the endpoint singularity. The integral in $h$ can be resolved numerically with weights and points by the Gauss-Legendre quadrature in absence of endpoint singularities. Gauss-Jacobi quadrature is recommended in that case.

>>>>

\begin{abstract}
In this work we apply the Ronen method to resolve the neutron transport in simple homogeneous problems. Slab, cylindrical and spherical geometries are studied. This method demands successive resolutions of the diffusion equation, where the local diffusion constants are modified in order to reproduce new estimates of the currents by a transport operator. The diffusion solver employs here finite differences and the transport-corrected currents are forced in the numerical scheme by means of drift terms, like in the CMFD technique. The solution is compared against reference results provided by the collision probability method. Boundary conditions are discussed introducing proper approximations to save the particle balance. Slow convergence is reported on the scalar flux, although the results match the reference solutions in the limit of fine meshes and far from the bare boundary.
%
%In this work we apply the Ronen method in the slab geometry filled with heterogeneous media. Neutron diffusion is solved by a nodal method on a coarse mesh, while reference solutions of the neutron distribution are provided by a discrete ordinate transport solver. The CMFD technique is used for both the implementation of the nodal solver and for the corrections of the diffusion coefficient given by the Ronen method. The presence of discontinuity factors for equivalencing the diffusion solution is investigated jointly with the Ronen method. Results are presented for different levels of material heterogeneity at the slab interfaces.
\end{abstract}
%
% SET RAGGED RIGHT MARGIN
%
\raggedright

@@

This idea was later used by Tomatis and Dall'Osso, who provided a numerical demonstration in a simple slab problem \cite{tomatis2011application}. Instead of updating the diffusion coefficient by the ratio of the current exchange and the flux gradient, as in Fick's law, they recurred to the scheme of the coarse mesh finite differences method (CMFD) for taking into account the new currents estimated by the integral transport operator in the diffusion solver.
%
%They proposed to update non-linearly the diffusion coefficients at the cell interfaces of the spatial mesh by the coarse mesh finite differences method (CMFD).
%
This technique, largely adopted in the literature of nodal methods \cite{smith1983nodal,lawrence1986progress}, can avoid indeterminate divisions in case of vanishing current and flux gradient. They tested this implementation in a bare slab with two-group cross sections homogenized in a realistic PWR assembly. It was observed that the Ronen method (RM) could drive the flux from diffusion towards the reference of the integral Boltzmann transport equation, regardless of the initial formulation used for the diffusion coefficient. Remarkably, the unphysical zero-flux boundary condition was not preventing to correct the flux near the boundary, that is within a few mean free paths. However, higher errors were noticed at the boundary with vacuum, slowly decreasing even after many iterations.% Their work was limited to the homogeneous slab, and resolved neutron diffusion on the same discretized spatial mesh used for the reference from neutron transport.
%
%In this work, we study the Ronen method with coarse meshes in order to approach the typical situation of core calculations, where homogenized cross sections are prepared in single fuel assemblies by lattice transport calculations. Nodal expansion methods are generally used with coarse meshes relying on polynomial-like or semi-analytical solutions \cite{lawrence1986progress}. They can also employ the discontinuity factors in their solving equations, thus yielding an equivalent form to the reference transport solution \cite{smith1986assembly}. Its implementation through the CMFD compensates the truncation errors caused by the approximate finite differences, and eases the global resolution to many local two-nodes problems. Of course, the simplified 1-D geometry let us disregard here all issues related to the transverse leakage approximation.
%
%Specifically, we compare hereafter the solutions obtained by the Ronen method and by the analytical nodal method, with the aim to present the first method as an on-line equivalence technique, farther capable to fulfill the neutron transport in its integral form on the coarse mesh. %We also use the discontinuity factors to enforce the physical equivalence between diffusion and transport, which is not preserved after the common flux-volume weighting homogenization of the cross sections. The implementation of both methods uses the CMFD, as shown in section \ref{sec:impl}. The basics of the analytical nodal method are reported in section \ref{sec:nd}, while the procedure to estimate the currents by the integral transport operator follows in section \ref{sec:ce}. New transport-corrected boundary conditions are derived in section \ref{sec:bc} to fix the bad performances observed in \cite{tomatis2011application}. The results on a few characteristic configurations are discussed in section \ref{sec:res}. Finally, the article ends with the conclusion in section \ref{sec:cncls}.

@@

The cross sections, as well as the diffusion coefficient, are usually available as volume-averaged data per cell in the mesh. A lattice transport computer code can prepare these data by homogenization. Once the scalar flux is known from the finite differences solver using the original diffusion coefficients, the integral expressions derived in section \ref{sec:ce} can be used to get new estimates of the current $J$ at the cell interfaces. %Alternatively, new current estimates can come from the two-nodes problem resolved by the nodal method in section \ref{sec:nd}.

@@

\subsection{Slab geometry}
\label{sec:slab}
The angular flux from the integral transport equation (with standard notation) is:
\begin{equation}
\varphi(x, \mu) = \varphi(x_b, \mu) e^{-\tau(x_b, x)/\mu} + \int_{x_0}^x {dx' \frac{q(x', \mu)}{\mu} e^{-\tau(x', x)/\mu}},
\label{eq:NTE}
\end{equation}
with the optical length $\tau(x_1, x_2) = \int_{x_1}^{x_2}{dx' \sigma_t(x')}$. Contrary to other formulations, we do not include the direction cosine of the neutron velocity $\mu$ in the definition of the optical length to ease the numerical integrations in the following. Eq. \ref{eq:NTE} holds for both positive directions $(\mu > 0)$ where $x_0 = x_{-1/2} = a$, and negative ones $(\mu > 0)$ where $x_0 = x_{I-1/2} = b$. The inclusion of the energy variable in Eq. \ref{eq:NTE} is straightforward; we use hereafter the multi-group theory, writing the source $q_g$ as\footnote{Curly brackets remind the possible use of the $k$-eigenvalue in case of vanishing external sources.}:
\begin{equation}
\label{eq:srcg}
q_g = \sum_{g'=1}^G{\left[
  \sum_{l=0}^\infty{ \frac{2l+1}{2}\sigma_{s,l, g' \rightarrow g}(x)P_l(\mu)\varphi_{l,g'}(x)
  + \frac{\chi_g}{2\{k\}} \nu\sigma_{f,g'}(x)\phi_{g'}(x)}\right] + S_g
                 },
\end{equation}
where the subscript $l$ refers to the moments of the expansions of the flux and of the scattering cross section on the Legendre polynomials $P_l$. Fission emission is isotropic, so that only the scalar flux $\phi$ being equal to the first moment $\varphi_0$ is retained for the fission term. Without the external source $S$, the multiplication factor $k$ is used as eigenvalue to avoid the only trivial vanishing solution. Since $q$ is computed by the results of the diffusion equation in the Ronen method, $l$ can only assume the values 0 and 1, with the second moment $\varphi_1$ equal to the net current $J$.% Hereafter we consider only isotropic sources.

Multiplication of Eq. \ref{eq:NTE} by $\mu$ and the integration over $[-1, 1]$ provides the expression for the current, see \cite{tomatis2011application}. The integration on $\mu$ yields integral exponential functions by a simple change of variable ($\mu = \pm 1/u$, according to the sign of $\mu$) \cite{AS1964handbook}:
\begin{equation*}
E_n(\tau) = \int_0^1{d\mu\, e^{-\tau / \mu} \mu^{n-2}} = \int_1^\infty{du\, e^{-\tau u}u^{-n}}, \: n \geq 0.
\end{equation*}
These special functions are then used to compute the current at the cell interfaces for all $i$:
\begin{equation}
\begin{split}
% J_g(x_{i+1/2}) &=
J_{g,i+1/2} &= \int_0^{ 1}{d\mu\, \mu \varphi_g(a, \mu) e^{-\tau_g(a, x_{i+1/2})/\mu}}
             - \int_0^{-1}{d\mu\, \mu \varphi_g(b, \mu) e^{-\tau_g(b, x_{i+1/2})/\mu}}\\
  & + \sum_{j=0}^{I-1} \left[  \sign(x_{i+1/2} - x_{j+1/2}) \frac{q_{0,g,j}}{2} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{2}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } \right.\\
  & \left. \qquad+ \frac{3}{2}q_{1,g,j} \int_{x_{j-1/2}}^{x_{j+1/2}}{dx' E_{3}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) }\right],
\end{split}
\label{eq:J_tr}
\end{equation}
where $q_{(.),g,j}$ are the moments of the souce $q_g$ from Eq. \ref{eq:srcg} averaged in the volume of the $j$-th cell:
\begin{equation*}
\label{eq:srclg}
\begin{split}
   q_{0,g,j} &= \sum_{g'}\left(\sigma_{s,0,g' \rightarrow g,j} +
                              \chi_g \nu\sigma_{f,g',j}
                        \right) \phi_{g',j},\\
   q_{1,g,j} &= \sum_{g'}\sigma_{s,1,g' \rightarrow g,j}J_{g',j}.
\end{split}
\end{equation*}
The current appearing in $q_{1,g,j}$, that is in case of linearly anisotropic scattering, is also to be considered as volume-averaged in the cell $j$. The optical lengths show the subscript $g$ because they are evaluated with the corresponding total cross section $\sigma_{t,g}$. For the property $E'_{n+1}(\ell) = -E_n(\ell)$, the spatial integrals of the integral exponential functions can be solved analytically by integrating on $\tau$:
\begin{equation*}
\begin{aligned}
\int_{x_{j-1/2}}^{x_{j+1/2}}{ dx' E_{n}\left(\lvert \tau_g(x', x_{i+1/2}) \rvert \right) } =
\frac{\sign(x_{i+1/2} - x_{j+1/2})}{\sigma_{t,g,j}} \cdot \\
\left[
  E_{n+1} \left(\lvert \tau_g(x_{j+1/2}, x_{i+1/2}) \rvert \right)
 -E_{n+1} \left(\lvert \tau_g(x_{j-1/2}, x_{i+1/2}) \rvert \right)
\right].
\end{aligned}
\end{equation*}
These quantities can be computed for $i > j$ only, because the others are simply opposite in sign.

@@

The projections of partial currents on the outward normal $\hat{n}$ to a given surface are obtained from integration of Eq. \ref{eq:ITE_xy} with the weight $| \hat{n} \cdot \mathbf{\Omega} | = \sin \theta | \hat{n} \cdot \mathbf{\Omega}_p |$ on $\hat{n} \cdot \mathbf{\Omega} \lessgtr 0$. Higher moments are obtained by the weights associated to the corresponding spherical harmonics. Integrations along the polar angle are resolved with the introduction of the Bickley-Naylor functions \cite{amos1983uniform}:
\begin{equation}
\text{Ki}_n(\tau) = \int_0^{\pi/2}{ d\theta \sin^{n-1} \theta \exp\left(-\frac{\tau}{\sin \theta}\right) },\: n \geq 0.
\label{eq:Kin}
\end{equation}
The properties used for the computation of these functions are available elsewhere \cite{lewis1984computational,hebert2009applied,amos1983algorithm}.

\begin{subequations}
After introducing these arguments for a bare cylinder, the current leaving the cylindrical surface at $\mathbf{r}$ (with direction $\hat{n}$) given by neutrons flying along $\mathbf{\Omega}_p$ is:
\begin{equation}
J^+(\mathbf{r}) = \frac{1}{2\pi}\int_{W^+}{
  d\omega\, | \hat{n} \cdot \mathbf{\Omega}_p | \,
  \int_0^S ds\, \text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r} - s \mathbf{\Omega}_p \right)\right]
  q(\mathbf{r} - s \mathbf{\Omega}_p)},
\label{eq:currp_cylinder1}
\end{equation}
with $W^+ = \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p > 0\right\}$. Using the surface element $dA' = s d\omega ds$ and $\mathbf{r}'_p = \mathbf{r} - s \mathbf{\Omega}_p$, it is also:
\begin{equation}
J^+(\mathbf{r}) = \int_{W^+ \times [0, S]}{
  dA'\, \hat{n} \cdot (\mathbf{r} - \mathbf{r}'_p) \,
  \frac{\text{Ki}_2\left[ \tau\left(\mathbf{r}, \mathbf{r}'_p \right)\right]}{2 \pi \lvert \mathbf{r} - \mathbf{r}'_p \rvert^2}
  q(\mathbf{r}'_p)}.
\label{eq:currp_cylinder2}
\end{equation}
\label{eq:currp_cylinder}
\end{subequations}
$J^-$ is obtained with integration over $w \in W^- =  \left\{\omega \mid \hat{n}\cdot \mathbf{\Omega}_p < 0\right\}$ instead.

@@

The corresponding form for the reduced escape probability for the incoming current is:
\begin{equation}
\label{eq:esc_prob_cyij-}
  \varepsilon^-_{i+1/2,j} = \frac{2}{r_{i+1/2}}
    \displaystyle ~\int_0^{r_{i+1/2}} { dh\, y
        \int_0^{\ell_j}{
            d\ell \text{Ki}_2 \left(\tau_{ji} + \sigma_j \ell \right)
        }
	},
\end{equation}
for $i<j$, whereas $\varepsilon^-_{i+1/2,j}\equiv 0$ for $i\geq j$ since in this case a neutron born in volume $j$ (or in volume $j=i$) cannot enter volume $i$ through its outer surface at radius $r_{i+1/2}$ without undergoing a collision.

Singularity of the integrand for the $h$-integration is noticed in the convex part while computing the collision probabilities \cite{hebert2009applied}; this notably happens at the outer radius of $j$-th ring where $\hat{n}$ is perpendicular to $\mathbf{\Omega}_p$. However, that integral can be regularized by a suitable change of variable. % $\varrho = \sqrt{r - h}$.
The calculation of the escape probabilities is not affected by this problem because of the factor $y$ that comes from the cosine of the angle described by the normal $\hat{n}$ and $\mathbf{\Omega}_p$, which cancels the endpoint singularity. The integral in $h$ can be resolved numerically with weights and points by the Gauss-Legendre quadrature in absence of endpoint singularities. Gauss-Jacobi quadrature is recommended otherwise.
